Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-22 Grassmannian_cells.tex
Abstract

\begin{document}
\begin{abstract}
  We prove that all quiver Grassmannians for preprojective representations of a generalized Kronecker quiver admit a cell decomposition.  
  In the process, we introduce a class of regular representations which arise as quotients of consecutive preprojective representations.
  Cell decompositions for quiver Grassmannians of these ``truncated preprojectives'' are also established as well as for all indecomposable representations in their Auslander-Reiten components.
  We also provide a natural combinatorial labeling for these cells using compatible pairs in a maximal Dyck path. 
\end{abstract}

@@

\tableofcontents
%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

A quiver Grassmannian is a projective variety attached to a fixed quiver representation which parametrizes subrepresentations of a fixed dimension.
In recent years, interest in quiver Grassmannians has grown considerably.
On the one hand, this is due to the fact that generating functions for the Euler characteristics of quiver Grassmannians of exceptional representations can be found as cluster variables.
On the other hand, they are clearly interesting on their own as they reveal many properties of the representation and its geometry.
Although it has been shown that every projective variety can be realized as a quiver Grassmannian -- which is maybe not surprising as the classification problem of indecomposable quiver representations is wild -- it turns out that very interesting phenomena arise when restricting to certain quivers or representations with certain properties.
For instance quiver Grassmannians attached to exceptional representations are smooth.
For tame quivers of types $\tilde A$ or $\tilde D$, it is known that every quiver Grassmannian attached to an indecomposable representation admits a cell decomposition into affine spaces.
It is conjectured that this is also true for exceptional representations of any quiver, in particular for preprojective and preinjective representations.

@@

As a benefit of this construction, we obtain a graph theoretic description of the non-empty cells.
More precisely, with every (truncated) preprojective representation we can associate a so-called $2$-quiver.
Essentially, such a quiver is obtained from a usual quiver by adding a collection of ``$2$-arrows'' between pairs of subquivers.
Now with every subset of the vertices we can associate a dimension vector.
If this subset is also strong successor closed, a condition which is easily verified in paractice, it corresponds to a cell and vice versa.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
Write $\rep Q$ for the hereditary abelian category of $\C$-representations of $Q$.

>>>>

\begin{document}
\begin{abstract}
  We prove that all quiver Grassmannians for preprojective representations of a generalized Kronecker quiver admit a cell decomposition into affine spaces.  
  In the process, we introduce a class of regular representations which arise as quotients of consecutive preprojective representations.
  Cell decompositions for quiver Grassmannians of these ``truncated preprojectives'' are also established. 
  We also provide two natural combinatorial labelings for these cells. On the one hand, the cells are in bijection with compatible pairs in a maximal Dyck path as predicted by the theory of cluster algebras. On the other hand, they are labeled by certain subsets of a so-called $2$-quiver attached to a preprojective representation. As there is a natural bijection between these two labelings this gives also a geometric explanation for the Dyck path combinatorics appearing in the theory of quiver Grassmannians. 
\end{abstract}

@@

\tableofcontents
%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\noindent A quiver Grassmannian is a projective variety attached to a fixed quiver representation which parametrizes subrepresentations of a fixed dimension.
In recent years, interest in quiver Grassmannians has grown considerably.
On the one hand, this is due to the fact that generating functions for the Euler characteristics of quiver Grassmannians of exceptional representations can be found as cluster variables.
On the other hand, they are clearly interesting on their own as they reveal many properties of the representation and its geometry.
Although it has been shown that every projective variety can be realized as a quiver Grassmannian -- which is maybe not surprising as the classification problem of indecomposable quiver representations is wild -- it turns out that very interesting phenomena arise when restricting to certain quivers or representations with certain properties.
For instance quiver Grassmannians attached to exceptional representations are smooth.
For tame quivers of types $\tilde A$ or $\tilde D$, it is known that every quiver Grassmannian attached to an indecomposable representation admits a cell decomposition into affine spaces.
It is conjectured that this is also true for exceptional representations of any quiver, in particular for preprojective and preinjective representations.

@@

As a benefit of this construction, we obtain a graph theoretic description of the non-empty cells.
More precisely, with every (truncated) preprojective representation we can associate a so-called $2$-quiver.
Essentially, such a quiver is obtained from a usual quiver by adding a collection of ``$2$-arrows'' between pairs of subquivers.
Now with every subset of the vertices we can associate a dimension vector.
If this subset is also strong successor closed, a condition which is easily verified in practice, it corresponds to a cell and vice versa.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.\sayT{Inserted finite-dimensional as we do not need infite-dimensional one (I am also not sure if we need to be careful at some points otherwise)}
